Accesible, realistic genome simulation with selection using \stdpopsim

### 2025-03-04 analysis2_manuscript.tex

\section*{Abstract}
    \label{abstract}
    Natural selection is a fundamental evolutionary force that shapes patterns of genetic variation across species. 
    However, simulating realistic models that incorporate both selection and complex demographic histories
    is challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and explore theoretical predictions.
    \stdpopsim is a tool for facilitating the simulation of a contiually expanding catalog of population genetic models.
    Here we present a major extension to the \stdpopsim{} framework that enables simulation of various modes
    of natural selection, including background selection, selective sweeps, and arbitrary distributions of fitness effects (DFE).
    This extension maintains \stdpopsim's core principles of reproducibility and standardization while adding support
    for species-specific genomic annotations and published DFE estimates. 
    We demonstrate the utility of this framework by benchmarking methods for demographic inference,
    DFE estimation, and selective sweep detection across different species and scenarios. 
    Our results demonstrate the robustness of demographic inference methods to selection on linked sites, 
    reveal the sensitivity of DFE inference methods to model assumptions, 
    and show how genomic features like recombination rate and the density of functional sequence influence power to detect selective sweeps.
    This extension to \stdpopsim{} provides a powerful new resource for the population genetics community
    to explore the interplay between selection and other evolutionary forces in a reproducible framework.

@@

    % stdpopsim
    In light of the complexities associated with estimating and simulating
    population genetic models, it is no surprise that
    a lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers \citep[e.g.,][]{ragsdale2020lessons}.
    This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
    While that is so, the original version of \stdpopsim did not include
    models of selection, which is a major limitation for more empirically motivated
    applications of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding processes such
    as adaptation\citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
    background selection on genetic diversity
    \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.
    Ideally, one would like
    to have a single, unified framework for simulating both neutral and
    non-neutral evolutionary processes, and to be able to compare the
    results of these simulations to empirical data in a manner that is
    both accessible to a wide range of researchers and highly reproducible. 
    Furthermore the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the size and density of functional elements, and
    non-equilibrium demographic histories. Ideally, one should be able
    to model all of these features simultaneously using estimates and annotations obtained
    from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.

@@

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for the curated addition of new
    species and models, and can use either
    msprime \citep{Baumdicker2022} or SLiM \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    %IG: edited here to add details on the spec of the DFE and Annotation classes
    %moved below the information on how they are used in simulation. Please check!
    To extend \stdpopsim's functionality to include models of selection, 
    we introduce two new classes
    %an important new class representing a specfic distribution of fitness effects
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where 
    each mutation type is associated with a dominance coefficient
    (currently limited to values between 0 and 1, inclusive),
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    Each \texttt{Annotation} object is typically associated with a collection of functional
    genomic elements whose coordinates are specified using a GFF3 file. %
    %    
    For example, say that we wish to implement a model of selection consistent with the
    DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonimous sites.
    with their proportions set to follow their expected
    prevelanece in exons. 
    The neutral mutation type is associted with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep{huber2018gene}
    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below).

    \begin{figure}[t!]
        \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
        \caption{\label{fig:schematic}
        \textbf{(A)} A schematic of the \stdpopsim catalog and the different components that
        can be used in simulations. 
        \textbf{(B)} The python API and \textbf{(C)} the command line interface.
        The user specifies a species, a portion of
        the genome to simulate, and optionally a genetic map, a model of
        demography, and a model of selection that itself is composed of a
        distribution of fitness effects (DFE) and a set of functional
        annotations. }
    \end{figure}

    As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on specices' publically available functional genomic elements 
    and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
    By using annotation objects, one can, for example, simulate populations where selected mutations (whose selection
    coefficients are drawn from a specified DFE) occur within annotation regions (e.g. exons), while all mutations
    occurring outside of these regions do not affect fitness.
    There are relatively few species with published DFE estimates in the literature.
    \stdpopsim currently has implemented DFEs for four species in the catalogue:
    \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
    humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While the DFEs are from a limited number of sepecies, these DFEs can be applied to other species
    in the catalog if and when it is suitable. 
    For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Furthermore, the user can specify a custom DFE and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have 
    a published DFE included in the catalog. This flexibility allows for a wide range of
    models of selection to be simulated. %, and for the results of these simulations to be
    %compared to empirical data or used as a null model for hypothesis testing.
    % DRS: commenting that stuff out because there are obviously other use cases that could be
    % listed here (e.g. ML/ABC, and if we don't know enough about a species to do that then
    % we probably have no business doing hypothesis testing either, which could be just as
    % misleading or maybe even more so), so the simplest solution is to cut this as it is
    % not needed here.

@@

    % current numbers of DFEs / species / etc
    % DRS: are we giving this a version number?
    At this release the \stdpopsim catalog comprises 24 species (i.e. genome representations)
    for which we have included 28 demographic models, 37 genetic maps, and 7 DFEs. %someone should double check the numbers
    Additions to the catalog are ongoing---we point the reader to our previously published 
    report detailing this effort \citep{lauterbur2023expanding}. We welcome contributions from the
    community and encourage those interested in contributing to visit https://popsim-consortium.github.io/stdpopsim-docs
    or contact one of the PopSim Consortium members for guidance.
    \igcomment{Mention releas number above, update the num of species and models and maybe also mention the number of annotation objects.}

@@

    \section*{Inference of $N_e(t)$ in the context of selection}
    One of the most common applications of population genetic inference is to estimate
    the effective population size over time, $N_e(t)$, from genetic data. This can be done
    using a variety of methods, including the sequential Markovian coalescent
    \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
    \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
    \drscomment{One could use dadi for that, but then we should also cite things like fastsimcoal and ABC, while explaining
    that users can use those methods to infer piecewise-constant N(t) histories as well with a prespecified number of epochs
    and priors around their params, but maybe that is too much of a digression here? I'm fine either way}
    through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
    as well as through identity by descent information \citep{santiago2020recent}.
    Since all of these methods assume neutral evolution, they are typically applied to genomic
    segments that are not expected to be directly affected by selection (e.g., by masking exons
    and conserved elements).
    However, natural selection acting on linked sites---even those that are masked out of the
    analysis---can bias estimates of $N_e(t)$
    away from true census population sizes by increasing the 
    rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}. 
    Using \stdpopsim, we can easily examine how different methods for inferring $N_e(t)$
    are influenced by selection on linked sites,
    which in the case of negative selection is commonly referred to as \emph{background selection}
    \citep{charlesworth1993effect,hudson1995deleterious}.

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
        \caption{
        \label{fig:human-demography}
        %citations aren't always allowed in captions. We should check with MBE.
        Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
        with and without purifying selection on exons. 
        Rows correspond to the three extant populations in the simulation: CHB, CEU, and YRI.
        Columns correspond to the four inference methods:    
        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
        Each plot depicts the inferred $N_e(t)$ on the three datasets simulated without selection (blue)
        and the three datasets simulated with the background effects of purifying selection on exons (orange),
        alongside the true values of $N_e$ used in simulation (black).
        }
    \end{figure}

@@

    Although the estimates produced by the four methods differ from one another, 
    for each method the estimates produced from data with selection (Figure
    \ref{fig:human-demography}; orange) and without selection (Figure
    \ref{fig:human-demography}; blue) are fairly similar.
    Thus, this particular parameterization of background selection does not
    appear to notably bias the estimates produced by these methods.
    As noted in previous studies that compared demography inference methods
    on neutral simulations \citep{adrion2020community},
    we see that each method has its strengths and weaknesses.
    Methods based on the sequential Markovian coalescent (\smcpp and \msmc)
    appear to overal produce the most accurate trajectories of $N_e(t)$
    in this setting, although they tend to under-estimate the population
    bottleneck associated with the population split.
    On the other hand, \stairway, which utilizes the site frequency spectrum,
    produces more noisy estimates,
    but it does not appear to under-estimate the population bottleneck.
    The inference of \gone, which utilizes identity by descent,
    is targeted toward demographic changes in more recent time frames ($\sim 200$ generations),
    explaining its noisy estimates for all but the most recent time periods.

@@

    To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
    We simulated 100 genomes under a two-epoch demographic model inferred for vaquita in \textcite{robinson2022critically},
    and a constant recombination rate across the genome (see \textbf{Methods}).
    In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
    annotated from the vaquita genome assembly.
    This DFE model implements a relationship between selection and dominance 
    in which very deleterious mutations (with selection coefficient $s<-0.1$) are fully recessive
    (dominance coefficient $h=0$) and weakly deleterious mutations ($s\geq -0.001$) are nearly
    additive ($h=0.4$; see \textbf{Methods}).
    Here, we see that the site-frequency-based method (\stairway) produces more accurate
    inference than the two methods based on the sequential Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
    Reagrdless of these differences, we see that background selection again does not
    appear to considerably influence the inference of any of the four methods for this species,
    which has a very different evolutionary history than humans.

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
        \drscomment{I think it would be good to increase the font size for the A,B,C panel legends and also
        draw a horizontal line above panel C so it is clear that all three plots at the bottom are part of C.}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons (see \textbf{Methods}).
        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulated datasets,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines.
        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets (averaging the inferred gamma parameters);
        white bars represent the distribution used in the simulation.
        }
        \label{fig:homsap-dfe.ooa}
    \end{figure}

    We start by examining DFE inference from human genomes simulated under an out-of-Africa demographic model
    with selection acting on exons (see \textbf{Methods}).
    In this setting, the DFE is inferred separately for each of the three simulated populations
    using three different DFE inference methods: \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
    and \grapes \citep{galtier2016adaptive}.
    In each case we note that the DFE is estimated from segregating sites only,
    without the use of substitution data, as \stdpopsim's models currently do not
    contain outgroup populations or species that can be used to estimate divergence.
    %IG: no need to map figures. Just describe results and cite figs
    %Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
    %while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
    %In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}).
    Note that while the estimation accuracy for the DFE parameters themselves (Figure \ref{fig:homsap-dfe.ooa}A-B) is useful,
    examining the difference between the actual and inferred distributions can be more helpful for showing how well a method
    has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C).
    The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters.
    As a result, \polydfe and \dadicli over-estimate the fraction of mutations under strong selection ($s<-0.1$),
    while \grapes considerably over-estimates the fraction of mutations under weak selection ($s\geq-0.0001$).
    % IG: commented this out to simplify the basic observation
    % and under-estimates the fraction of mutations with $s<-0.01$.

    Population history appears to influence different methods in slightly different ways.
    For instance, \dadicli and \grapes produce more accurate estimates from the YRI genomes 
    (relative to the other two populations), whereas \polydfe
    produces its most accurate estimate from the CHB genomes.
    To factor out the influence of demographic changes, we also analyzed genomes generated using similar
    simulations but with no population size changes (Figure \ref{fig:homsap-dfe.constant}).
    In this simpler setting, \polydfe overlapped the true mean selection coefficient and shape parameter, whereas 
    the other methods do not. 
    \grapes exhibits similar biases without population size changes as in the simulations with more complex demography,
    but with milder under-estimation of the shape parameter, but \dadicli shows an inverse pattern
    with overestimated shape parameter and slightly underestimated mean.

    \begin{figure}[b!]
        \centering
        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
        \caption{
        \label{fig:vaquita-dfe}
        Inferred versus simulated DFEs from simulated vaquita genomes.
        Simulations were performed using a two-epoch model of vaquita porpoise demography with a gamma-distributed DFE
        acting on exons with a relationship between the selection coefficient ($s$) and dominance coefficient ($h$) (see \textbf{Methods}).
        The DFE is inferred by analyzing all simulated genomes jointly by one of three different methods:
        \grapes, \polydfe, and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulated datasets,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines. 
        (C) Binned distribution of $2hs$ implied by the simulated DFE compared with
        the distribution of $s$ implied by the average DFE inferred for each method from the three simulated datasets (averaging the inferred gamma parameters).
        The distribution of $2 h s$ is multimodal because of the simulated relationship between $h$ and $s$ (see text).
        }
    \end{figure}

>>>>

\section*{Abstract}
    \label{abstract}
    Natural selection is a fundamental evolutionary force that shapes patterns of genetic variation across species. 
    However, simulating realistic models that incorporate both selection and complex demographic histories
    is challenging, limiting our ability to benchmark statistical methods aimed at detecting selection and explore theoretical predictions.
    \stdpopsim is a tool for facilitating the simulation of a contiually expanding catalog of population genetic models.
    Here we present a major extension to the \stdpopsim{} framework that enables simulation of various modes
    of natural selection, including background selection, selective sweeps, and arbitrary distributions of fitness effects (DFE).
    This extension maintains \stdpopsim's core principles of reproducibility and standardization while adding support
    for species-specific genomic annotations and published DFE estimates. 
    We demonstrate the utility of this framework by benchmarking methods for demographic inference,
    DFE estimation, and selective sweep detection across different species and scenarios. 
    Our results highlight how selection can bias demographic inference, 
    reveal the sensitivity of DFE inference methods to model assumptions, 
    and show how genomic features like recombination rate and the density of functional sequence influence power to detect selective sweeps.
    This extension to \stdpopsim{} provides a powerful new resource for the population genetics community
    to explore the interplay between selection and other evolutionary forces in a reproducible framework.

@@

    % stdpopsim
    In light of the complexities associated with estimating and simulating
    population genetic models, it is no surprise that
    a lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers \citep[e.g.,][]{ragsdale2020lessons}.
    This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
    While that is so, the original version of \stdpopsim did not include
    models of selection, which is a major limitation for more empirically motivated
    applications of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding processes such
    as adaptation\citep[e.g.][]{thornton2019polygenic,hartfield2024polygenic}, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
    background selection on genetic diversity
    \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.
    Ideally, one would like
    to have a single, unified framework for simulating both neutral and
    non-neutral evolutionary processes, and to be able to compare the
    results of these simulations to empirical data in a manner that is
    both accessible to a wide range of researchers and highly reproducible. 
    Furthermore the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the size and density of functional elements, and
    non-equilibrium demographic histories. Ideally, one should be able
    to model all of these features simultaneously using estimates/annotations obtained
    from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.

@@

    % implementing selection
    \stdpopsim previously provided a wide range of models of neutral
    demographic history from roughly two dozen species \citep{lauterbur2023expanding}.
    This was accomplished through the use of a standardized interface
    and data structure that allows for the curated addition of new
    species and models, and can use either
    msprime \citep{Baumdicker2022} or SLiM \citep{haller2019slim}
    as the backend engine for efficient population genetic simulation.
    \igcomment{edited here to add details on the spec of the DFE and Annotation classes
    moved below the information on how they are used in simulation. Please check!}
    To extend \stdpopsim's functionality to include models of selection, 
    we introduce two new classes
    %an important new class representing a specfic distribution of fitness effects
    to the \stdpopsim API: the \texttt{DFE} class and the \texttt{Annotation} class.
    The \texttt{DFE} class provides an interface for specifying
    a distribution of the fitness effects of new mutations.
    Each \texttt{DFE} object specifies a list of mutation types and
    their proportions, where 
    each mutation type is associated with a dominance coefficient
    (currently limited to values between 0 and 1, inclusive),
    the distribution of selection coefficients (e.g., fixed-value, normally-distributed, gamma-distributed),
    and the parameters of this distribution.
    Each \texttt{Annotation} object is typically associated with a collection of functional
    genomic elements whose coordinates are specified using a GFF3 file. %
    %    
    For example, say that we wish to implement a model of selection consistent with the
    DFE inferred by Kim et al. (\citeyear{kim2017inference}) for exons in the human genome.
    The associated \texttt{DFE} object is defined with two mutation types,
    neutral and deleterious, corresponding to synonymous and non-synonimous sites.
    with their proportions set to follow their expected
    prevelanece in exons. 
    The neutral mutation type is associted with a fixed selection coefficient (0),
    and the deleterious mutation type is associated with a dominance coefficient of $0.5$ and
    a gamma-distributed selection coefficient
    with mean and shape as inferred by Kim et al. (\citeyear{kim2017inference}).
    \texttt{DFE} objects can also model a relationship between the selection and dominance coefficients \citep{huber2018gene}
    by defining ranges of selection coefficients and associating each range with a different dominance coefficient (see below).

    \begin{figure}[t!]
        \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
        \caption{\label{fig:schematic}
        \textbf{(A)} A schematic of the \stdpopsim catalog and the different components that
        can be used in simulations. 
        \textbf{(B)} The python API and \textbf{(C)} the command line interface.
        The user specifies a species, a portion of
        the genome to simulate, and optionally a genetic map, a model of
        demography, and a model of selection that itself is composed of a
        distribution of fitness effects (DFE) and a set of functional
        annotations. \rngcomment{I find the images too small to read the text comfortably.}
        \igcomment{I changed this to full width. We might want to reduce font size in the panel titles.}}
    \end{figure}

    \igcomment{modified this paragraph to focus on the catalog and moved simulation details to next paragraph}
    As with other aspects of \stdpopsim, we intend the selection models to be as biologically accurate as possible.
    Thus, the \stdpopsim catalog provides \texttt{Annotation} objects based on specices' publically available functional genomic elements 
    and \texttt{DFE} objects based on published DFE estimates (Figure \ref{fig:schematic}A).
    By using annotation objects, one can, for example, simulate populations where selected mutations (whose selection
    coefficients are drawn from a specified DFE) occur within annotation regions (e.g. exons), while all mutations
    occurring outside of these regions are fitness neutral.
    There are relatively few species with published DFE estimates in the literature.
    \stdpopsim currently has implemented DFEs for four species in the catalogue:
    \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
    humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While the DFEs are from a limited number of sepecies, these DFEs can be applied to other species
    in the catalog if and when it is suitable. 
    For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Furthermore, the user can specify a custom DFE and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have 
    a published DFE included in the catalog. This flexibility allows for a wide range of
    models of selection to be simulated. %, and for the results of these simulations to be
    %compared to empirical data or used as a null model for hypothesis testing.
    % DRS: commenting that stuff out because there are obviously other use cases that could be
    % listed here (e.g. ML/ABC, and if we don't know enough about a species to do that then
    % we probably have no business doing hypothesis testing either, which could be just as
    % misleading or maybe even more so), so the simplest solution is to cut this as it is
    % not needed here.

@@

    % current numbers of DFEs / species / etc
    % DRS: are we giving this a version number?
    At this release the \stdpopsim catalog comprises 24 species (i.e. genome representations)
    for which we have included 28 demographic models, 37 genetic maps, and 7 DFEs. %someone should double check the numbers
    \igcomment{do we want to also mention the number of annotation objects?}
    Additions to the catalog are ongoing---we point the reader to our previously published 
    report detailing this effort \citep{lauterbur2023expanding}. We welcome contributions from the
    community and encourage those interested in contributing to visit https://popsim-consortium.github.io/stdpopsim-docs
    or contact one of the PopSim Consortium members for guidance.

@@

    \section*{Inference of $N_e(t)$ in the context of selection}
    One of the most common applications of population genetic inference is to estimate
    the effective population size over time, $N_e(t)$, from genetic data. This can be done
    using a variety of methods, including the sequentially Markovian coalescent
    \citep{li2011inference,Schiffels2020,terhorst2017robust}, \kelcomment{mabye also cite Gutenkunst 2009 Dadi paper here} 
    \igcomment{I'm not sure it's all that relevant, since it doesn't strictly infer N(t)}
    \drscomment{One could use dadi for that, but then we should also cite things like fastsimcoal and ABC, while explaining
    that users can use those methods to infer piecewise-constant N(t) histories as well with a prespecified number of epochs
    and priors around their params, but maybe that is too much of a digression here? I'm fine either way}
    through use of the site frequency spectrum (SFS) \citep{liu2020stairway},
    as well as through identity by descent information \citep{santiago2020recent}.
    Since all of these methods assume neutral evolution, they are typically applied to genomic
    segments that are not expected to be directly affected by selection (e.g., by masking exons
    and conserved elements).
    However, natural selection acting on linked sites---even those that are masked out of the
    analysis---can bias estimates of $N_e(t)$
    away from true census population sizes by increasing the 
    rate of coalescence in the genome \citep[e.g.][]{schrider2016effects}. 
    Using \stdpopsim, we can easily examine the effect of selection on linked sites,
    which in the case of negative selection is commonly referred to as \emph{background selection} \citep{charlesworth1993effect,hudson1995deleterious}, 
    on different methods for inferring $N_e(t)$.

@@

    \begin{figure}[b!]
        \centering
        %IG: NEED TO LINK NEW FIG HERE
        %\includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
        \igcomment{NEED TO ADD NEW VERSION OF FIG THAT SILAS GENERATED}
        \caption{
        \label{fig:human-demography}
        %citations aren't always allowed in captions. We should check with MBE.
        Inference of $N_e(t)$ from human genomes simulated under an out-of-Africa demographic model \citep{ragsdale2019models}
        with and without purifying selection on exons. 
        Rows correspond to the three extant populations in the simulation: CHB, CEU, and YRI.
        Columns correspond to the four inference methods:    
        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.
        Each plot depicts the inferred $N_e(t)$ on the three datasets simulated without selection (blue)
        and the three datasets simulated with the background effects of purifying selection on exons (orange),
        alongside the true values of $N_e$ used in simulation (black).
        }
    \end{figure}

    \igcomment{Rewrote this paragraph based on new figs.}
    %
    Although the estimates produced by the four methods differ from one another, 
    for each method the estimates produced from data with selection (Figure
    \ref{fig:human-demography}; orange) and without selection (Figure
    \ref{fig:human-demography}; blue) are fairly similar.
    Thus, this particular parameterization of background selection does not
    appear to notably bias the estimates produced by these methods.
    As noted in previous studies that compared demography inference methods
    on neutral simulations \citep{adrion2020community},
    we see that each method has its strengths and weaknesses.
    Methods based on the sequenial Markovian coalescent (\smcpp and \msmc)
    appear to overal produce the most accurate trajectories of $N_e(t)$
    in this setting, although they tend to under-estimate the population
    bottleneck associated with the population split.
    On the other hand, \stairway, which utilizes the site frequency spectrum,
    produces more noisy estimates,
    but it does not appear to under-estimate the population bottleneck.
    The inference of \gone, which utilizes identity by descent,
    is targeted toward demographic changes in more recent time frames ($\sim 200$ generations),
    explaining its noisy estimates for all but the most recent time periods.

@@

    To expand these observations to additional species and to highlight the ease of comparison enabled by \stdpopsim,
    we performed a similar benchmark using the vaquita porpoise \textit{Phocoena sinus}.
    We simulated 100 genomes under a two-epoch demographic model inferred for vaquita in \textcite{robinson2022critically},
    and a constant recombination rate across the genome (see \textbf{Methods}).
    In simulations with selection, we applied the DFE inferred by \textcite{robinson2022critically} to exons
    annotated from the vaquita genome assembly.
    This DFE model implements a relationship between selection and dominance 
    in which very deleterious mutations (with selection coefficient $s<-0.1$) are fully recessive
    (dominance coefficient $h=0$) and weakly deleterious mutations ($s\geq -0.001$) are nearly
    additive ($h=0.4$; see \textbf{Methods}).
    Here, we see that the site-frequency-based method (\stairway) produces more accurate
    inference than the two methods based on the sequenctioal Markovian coalescent (\smcpp and \msmc; see Figure \ref{fig:vaquita-demography}).
    Reagrdless of these differences, we see that background selection again does not
    appear to considerably influence the inference of any of the four methods for this species,
    which has a very different evolutionary history than humans.

@@

    \begin{figure}[b!]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_discrete_DFE}
        \drscomment{I think it would be good to increase the font size for the A,B,C panel legends and also
        draw a horizontal line above panel C so it is clear that all three plots at the bottom are part of C.}
        \caption{Inferred versus simulated DFEs from simulated human genomes.
        Simulations were performed using a human out-of-Africa demographic model with a gamma-distributed DFE
        acting on exons (see \textbf{Methods}).
        DFE is inferred separately for each of the three extant populations (CEU, CHB, and YRI)
        by three different methods: \grapes, \polydfe , and \dadicli.
        (A-B) Mean absolute value of selection coefficient ($\lvert E(s) \rvert $) and shape parameter are
        shown for each DFE inferred from all three simulated datasets,
        with median values marked by horizontal bars
        and simulated values represented by dashed horizontal lines.
        (C) Binned distribution of $s$ implied by the average DFE inferred from the three simulated datasets;
        white bars represent the distribution used in the simulation.
        }
        \label{fig:homsap-dfe.ooa}
    \end{figure}

    \igcomment{I start with the OOA model and then cite the constant-size model as supplementary.}
    We start by examining DFE inference from human genomes simulated under an out-of-Africa demographic model
    with selection acting on exons (see \textbf{Methods}).
    In this setting, the DFE is inferred separately for each of the three simulated populations
    using three different DFE inference methods: \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
    and \grapes \citep{galtier2016adaptive}.
    In each case we note that the DFE is estimated from segregating sites only,
    without the use of substitution data, as \stdpopsim's models currently do not
    contain outgroup populations or species that can be used to estimate divergence.
    %IG: no need to map figures. Just describe results and cite figs
    %Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
    %while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
    %In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
    Overall, inferred DFEs varied across populations and inference methods (Figure \ref{fig:homsap-dfe.ooa}).
    Note that while the estimation accuracy for the DFE parameters themselves (Figure \ref{fig:homsap-dfe.ooa}A,B) is useful,
    examining the difference between the actual and inferred distributions can be more helpful for showing how well a method
    has recovered the true DFE (Figure \ref{fig:homsap-dfe.ooa}C)
    The distributions of $s$ inferred by \polydfe and \dadicli had higher mean values (in absolute value)
    but smaller shape parameters, while \grapes inferred distributions with lower mean values and shape parameters (Figure \ref{fig:homsap-dfe.ooa}A-B).
    As a result, \polydfe and \dadicli over-estimate the fraction of mutations under strong selection ($s<-0.1$; Figure \ref{fig:homsap-dfe.ooa}C),
    while \grapes under-estimates the fraction of mutations with $-1<s<-0.01$ and considerably over-estimates the fraction of
    mutations under weak selection ($s\geq-0.001$).
    Population history appears to influence different methods in slightly different ways.
    For instance, \dadicli and \grapes produce more accurate estimates from the YRI genomes 
    (relative to the other two populations), whereas \polydfe
    produces its most accurate estimate from the CHB genomes.
    To factor out the influence of demographic changes, we also analyzed genomes generated using similar
    simulations but with no population size changes (Figure \ref{fig:homsap-dfe.constant}).
    In this simpler setting, \polydfe overlapped the true mean selection coefficient and shape parameter, whereas 
    the other methods do not. 
    \grapes exhibits similar biases without population size changes as in the simulations with more complex demography,
    but with milder under-estimation of the shape parameter, but \dadicli shows an inverse pattern
    with overestimated shape parameter and slightly underestimated mean. 
\kelcomment{I'm a little surprised by the non-great performance of Dadi-cli in here....I know that folks have spent a long time working on this analysis and debugging it, so I don't want be negative or delay things. That said, I think it would be good to plot the empirical (in the case simulated) SFS along with the SFSs from the MLE models. I would also do this for the demogrpahic models as well as the DFE models}
\igcomment{I think that this is better conveyed now in the added histograms for $s$ in panel C}
